Robust Color Maps That Work for Most Audiences (Including the U.S. President)

### 2020-09-22 pkg/colorspace/paper/ZAMG2020/slides.Rnw
Adding slides for ZAMG internal presentation, September 2020.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/colorspace@555 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
%\usetheme[nototalframenumber, noslidenumber, license, url]{uibk}
\usetheme[license, url]{uibk}
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{colorspace.png}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{colorspace.png}%
      ]%
}
\headerimage{1}

\title{A Toolbox for Manipulating and Assessing Color Palettes}
\subtitle{\phantom{.}\\Achim Zeileis, Jason C. Fisher, Kurt Hornik, Ross Ihaka, Claire D. McWhite, Paul Murrell, Reto Stauffer, Claus O. Wilke}
\author{\vspace*{-6.3cm}\hspace*{12.5cm}\includegraphics[width=1.6cm]{palmfooter.png}\\[5.5cm]} %% -6.83 with "for Statistical Graphics"
\URL{http://colorspace.R-Forge.R-project.org/}

@@

<<preliminaries>>=
library("colorspace")
library("ggplot2")
theme_set(theme_minimal())
@

@@

\section{colorspace: A Toolbox for Manipulating and Assessing Color Palettes}

\subsection{Motivation}

\begin{frame}
\frametitle{Motivation}

@@

\begin{minipage}{0.54\textwidth}
\only<1,4>{\includegraphics[width=\textwidth]{../../vignettes/screenshots/influenza-rainbow.png}}%
\only<2>{\includegraphics[width=\textwidth]{../../vignettes/screenshots/influenza-rainbow-deutan.png}}%
\only<3>{\includegraphics[width=\textwidth]{../../vignettes/screenshots/influenza-rainbow-gray.png}}%
\only<5>{\includegraphics[width=\textwidth]{../../vignettes/screenshots/influenza-purpleyellow.png}}%
\only<6>{\includegraphics[width=\textwidth]{../../vignettes/screenshots/influenza-purpleyellow-deutan.png}}%
\only<7>{\includegraphics[width=\textwidth]{../../vignettes/screenshots/influenza-purpleyellow-gray.png}}
\end{minipage}
\hspace*{0.6cm}
\begin{minipage}{0.4\textwidth}

\textbf{Map:} Influenza severity in Germany (week 8, 2019).\\

\textbf{Source:} Arbeitsgemeinschaft Influenza, Robert-Koch-Institut.\\

\textbf{Reported in:} SPIEGEL Online, Tagesschau, \dots\\

\visible<2->{\textbf{Problems:} Color vision deficiency.}
\visible<3->{Grayscale.}
\visible<4->{Flashy.\\}

\visible<5->{\textbf{Alternative:} HCL-based sequential palette.}

\end{minipage}

@@

\begin{frame}
\frametitle{Motivation}

\textbf{Need tools for:}
\begin{itemize}
  \item Construction of palettes with better perceptual properties.
  \item Assessment of color palettes.
  \item Manipulation of colors.
\end{itemize}

\begin{minipage}{0.5\textwidth}
\includegraphics[width=\textwidth]{../../vignettes/screenshots/bobross.jpg}
\end{minipage}
\hspace*{0.6cm}
\begin{minipage}{0.36\textwidth}
\vspace*{-4cm}
Because Bob Ross would not approve of this!
\end{minipage}

@@

\begin{frame}
\frametitle{Color spaces}

\textbf{Origin of the package:} Convert colors between various three-dimensional representations of color.

\textbf{In particular:} From the perceptually-based HCL (Hue-Chroma-Luminance) to standard Red-Green-Blue (sRGB, and corresponding hex codes) space.

\setkeys{Gin}{width=0.8\textwidth}
\begin{center}
<<color-spaces, fig = TRUE, width = 9.5, height = 4.5>>=
par(mar = rep(0, 4))

d <- data.frame(
  x = c(1, 1, 2, 2, 3, 4, 5, 5, 5, 6),
  y = c(1, 3, 1, 3, 2, 2, 1, 3, 2, 2),
  lab = c("polarLAB", "polarLUV\n(= HCL)", "LAB", "LUV", "XYZ", "RGB", "HLS", "HSV", "sRGB", "hex"),
  stringsAsFactors = FALSE
)

plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
  xlim = c(0.7, 6.3), ylim = c(0.4, 3.6))
lines(y ~ x, data = d[-c(1, 3, 7, 8),])
lines(y ~ x, data = d[c(1, 3, 5),])
lines(y ~ x, data = d[c(6, 7, 9),])
lines(y ~ x, data = d[c(6, 8, 9),])
points(y ~ x, data = d[c(1, 2, 7, 8), ], pch = 21, bg = "lightgray", cex = 12)
points(y ~ x, data = d[c(3, 4, 5, 6, 9), ], pch = 22, bg = "lightgray", cex = 12)
points(y ~ x, data = d[10, ], pch = 23, bg = "white", cex = 12)

text(d$x, d$y, d$lab)
text(3.5, 2, "white point\n= D65", cex = 0.7, pos = 3)
text(4.56, 2, "gamma\n(= 2.4)", cex = 0.7, pos = 3)
@
\end{center}

@@

\begin{frame}
\frametitle{HCL vs.~RGB}

@@

@@

\setkeys{Gin}{width=0.44\textwidth}
<<hcl-rainbow, fig = TRUE, width = 5.2, height = 6>>=
specplot(rainbow_hcl(99, c = 60, l = 75), rgb = TRUE)
@
\hspace*{0.8cm}
\pause
<<rgb-rainbow, fig = TRUE, width = 5.2, height = 6>>=
specplot(rainbow(95), rgb = TRUE)
@

@@

\only<1>{
\textbf{Sequential:} Luminance contrast is crucial (dark to light or vice versa).
}
\only<2>{
\textbf{Blues 2:} Single hue. Decreasing chroma with increasing luminance.
}
\only<3>{
\textbf{Blues 3:} Single hue. Triangular chroma to achieve higher luminance contrast.
}
\only<4>{
\textbf{Blues:} Multi hue. Triangular chroma. High luminance contrast.
}

\begin{minipage}[t]{0.49\textwidth}
\vspace*{0.46cm}

\setkeys{Gin}{width=\textwidth}
<<blues, fig = TRUE, width = 4, height = 2.2>>=
swatchplot(
  "Blues 2" = sequential_hcl(7, "Blues 2"),
  "Blues 3" = sequential_hcl(7, "Blues 3"),
  "Blues"   = sequential_hcl(7, "Blues")
)
@
\end{minipage}
\only<2>{%
\begin{minipage}[t]{0.49\textwidth}
\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<blues2-spectrum, fig = TRUE, width = 5.2, height = 5.2>>=
specplot(sequential_hcl(7, "Blues 2"))
@
\end{minipage}%
}%
\only<3>{%
\begin{minipage}[t]{0.49\textwidth}
\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<blues3-spectrum, fig = TRUE, width = 5.2, height = 5.2>>=
specplot(sequential_hcl(7, "Blues 3"))
@
\end{minipage}
}%
\only<4>{%
\begin{minipage}[t]{0.49\textwidth}
\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<blues-spectrum, fig = TRUE, width = 5.2, height = 5.2>>=
specplot(sequential_hcl(7, "Blues"))
@
\end{minipage}
}

@@

\subsection{Statistical graphics}

\begin{frame}[fragile]
\frametitle{Statistical graphics}

\textbf{Base:}

\begin{itemize}
  \item HCL palette functions return hex color vector.
  \item Typically passed to \code{col = } argument of base plotting functions.
\end{itemize}

\textbf{ggplot2:}

\begin{itemize}
  \item Scales of type \code{scale_<aesthetic>_<datatype>_<colorscale>()}.
  \item \code{<aesthetic>} is \code{fill} or \code{color}/\code{colour}.
  \item \code{<datatype>} is \code{discrete} or \code{continuous}.
  \item \code{<colorscale>} is \code{qualitative}, \code{sequential}, \code{diverging}, or \code{divergingx}.
\end{itemize}

\begin{frame}[fragile]
\frametitle{Statistical graphics: Base}

<<base-graphics, eval = FALSE, echo = TRUE>>=
q4 <- qualitative_hcl(4, palette = "Dark 3")
plot(log(EuStockMarkets), plot.type = "single", col = q4, lwd = 2)
legend("topleft", colnames(EuStockMarkets), col = q4, lwd = 3, bty = "n")
@

\vspace*{-0.6cm}

\setkeys{Gin}{width=0.8\textwidth}
<<base-graphics-plot, fig = TRUE, width = 10, height = 6>>=
<<base-graphics>>
@

\begin{frame}[fragile]
\frametitle{Statistical graphics: Base}

<<base-graphics2, eval = FALSE, echo = TRUE>>=
ttnc <- margin.table(Titanic, c(1, 4))[, 2:1]
spineplot(ttnc, col = sequential_hcl(2, palette = "Purples 3"))
@

\vspace*{-0.6cm}

\setkeys{Gin}{width=0.8\textwidth}
<<base-graphics2-plot, fig = TRUE, width = 10, height = 6>>=
<<base-graphics2>>
@

\begin{frame}[fragile]
\frametitle{Statistical graphics: ggplot2}

\setkeys{Gin}{width=0.8\textwidth}
<<ggplot2-graphics, fig = TRUE, width = 10, height = 5.5, echo = TRUE>>=
library("ggplot2")
ggplot(iris, aes(x = Sepal.Length, fill = Species)) + geom_density(alpha = 0.6) +
  scale_fill_discrete_qualitative(palette = "Dark 3")
@

\begin{frame}[fragile]
\frametitle{Statistical graphics: ggplot2}

\setkeys{Gin}{width=0.8\textwidth}
<<ggplot2-graphics2, fig = TRUE, width = 10, height = 5.5, echo = TRUE>>=
dsamp <- diamonds[1 + 1:1000 * 50, ]
ggplot(dsamp, aes(carat, price, color = cut)) + geom_point() +
  scale_color_discrete_sequential(palette = "Purples 3", nmax = 6, order = 2:6)
@

@@

@@

\begin{frame}
\frametitle{Base R\only<2>{: Why you might not need our package after all}}

@@

@@

\begin{frame}
\frametitle{References}

@@

>>>>

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
%\usetheme[nototalframenumber, noslidenumber, license, url]{uibk}
\usetheme[nototalframenumber, license, url]{uibk}
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{colorspace.png}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{colorspace.png}%
      ]%
}
\headerimage{1}

\title{Robust Color Maps That Work for Most Audiences (Including the U.S. President)}
\subtitle{Reto Stauffer, Achim Zeileis}
\author{\vspace*{-6.1cm}\hspace*{12.5cm}\includegraphics[width=1.6cm]{palmfooter.png}\\[5.5cm]\vspace{-1.5em}} %% -6.83 with "for Statistical Graphics"
\URL{http://hclwizard.org/}%

@@

<<preliminaries>>=
library("colorspace")
@

@@

\section{colorspace: A Toolbox for Manipulating and Assessing Color Palettes} %

\subsection{Motivation}%

% ---------------------------------------------------------------------
\begin{frame}
\frametitle{Motivation}

\vspace*{2em}

\begin{minipage}{0.46\textwidth}
    \includegraphics[width=\textwidth]{images/sharpie-small.png}
\end{minipage}
\pause
\hfill
\begin{minipage}{0.51\textwidth}
    \includegraphics[width=\textwidth]{images/tweet.png}
\end{minipage}

    \vspace*{0.5em}
    \begin{footnotesize}
    Projected track and wind speed forecast of hurricane Dorian.
    Screenshot of a video released by the White House (Sep. 4, 2019),
    tweet by the U.S. president (Sep. 5, 2019).
    \end{footnotesize}

%% ---------------------------------------------------------------------
%\begin{frame}
%\frametitle{Motivation}
%
%\vspace*{-0.3cm}
%
%\begin{minipage}{0.54\textwidth}
%    \only<1>{\includegraphics[width=\textwidth]{images/dorian-rainbow.png}}%
%    \only<2-3>{\includegraphics[width=\textwidth]{images/dorian-rainbow-arrow.png}}%
%    \only<4>{\includegraphics[width=\textwidth]{images/dorian-rainbow-deutan.png}}%
%    \only<5>{\includegraphics[width=\textwidth]{images/dorian-rainbow-gray.png}}%
%    \only<6>{\includegraphics[width=\textwidth]{images/dorian-hclrainbow.png}}%
%    \only<7>{\includegraphics[width=\textwidth]{images/dorian-hclrainbow-gray.png}}
%    \only<8>{\includegraphics[width=\textwidth]{images/dorian-hclrainbow-deutan.png}}
%\end{minipage}
%\hspace*{0.6cm}
%\begin{minipage}{0.4\textwidth}
%
%    \textbf{Map:} Probability of exceeding wind speeds of 63\,km\,h$^{-1}$,
%    Aug~30 -- Sep~04, 2019.
%
%    \medskip
%    \textbf{Source:} NOAA (\href{https://noaa.gov}{noaa.gov})
%
%    \medskip
%    \visible<3->{\textbf{Problems:} Flashy.}
%    \visible<4->{Color vision deficiency.}
%    \visible<5->{Grayscale.}
%    
%    \medskip
%    \visible<6->{\textbf{Alternative:} HCL-based sequential palette.}
%
%\end{minipage}
%
%\end{frame}

% ---------------------------------------------------------------------
\begin{frame}
    \frametitle{\only<1-4>{Problem}\only<5->{Solution}}

@@

    \begin{center}
        \only<1>{\includegraphics[width=.55\textwidth]{images/dorian-rainbow.png}}%
        \only<2>{\includegraphics[width=.55\textwidth]{images/dorian-rainbow-arrow.png}}%
        \only<3>{\includegraphics[width=.55\textwidth]{images/dorian-rainbow-deutan.png}}%
        \only<4>{\includegraphics[width=.55\textwidth]{images/dorian-rainbow-gray.png}}%
        \only<5>{\includegraphics[width=.55\textwidth]{images/dorian-hclrainbow.png}}%
        \only<6>{\includegraphics[width=.55\textwidth]{images/dorian-hclrainbow-gray.png}}
        \only<7>{\includegraphics[width=.55\textwidth]{images/dorian-hclrainbow-deutan.png}}
    \end{center}

    \vspace*{-0.5em}
    \begin{footnotesize} 
    Probability of exceeding wind speeds of 39\,mph (63\,km\,h$^{-1}$),
    Aug~30 -- Sep~04, 2019 (\href{https://noaa.gov}{noaa.gov}).
    \end{footnotesize}

@@

%% ---------------------------------------------------------------------
%\begin{frame}
%\frametitle{Motivation}
%
%\textbf{Need tools for:}
%\begin{itemize}
%  \item Construction of palettes with better perceptual properties.
%  \item Assessment of color palettes.
%  \item Manipulation of colors.
%\end{itemize}
%
%\end{frame}

@@

%% ---------------------------------------------------------------------
%\begin{frame}
%\frametitle{Color spaces}
%
%\textbf{Origin of the package:} Convert colors between various three-dimensional representations of color.
%
%\medskip
%
%\textbf{In particular:} From the perceptually-based HCL (Hue-Chroma-Luminance) to standard Red-Green-Blue (sRGB, and corresponding hex codes) space.
%
%\setkeys{Gin}{width=0.8\textwidth}
%\begin{center}
%<<color-spaces, fig = TRUE, width = 9.5, height = 4.5>>=
%par(mar = rep(0, 4))
%
%d <- data.frame(
%  x = c(1, 1, 2, 2, 3, 4, 5, 5, 5, 6),
%  y = c(1, 3, 1, 3, 2, 2, 1, 3, 2, 2),
%  lab = c("polarLAB", "polarLUV\n(= HCL)", "LAB", "LUV", "XYZ", "RGB", "HLS", "HSV", "sRGB", "hex"),
%  stringsAsFactors = FALSE
%)
%
%plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
%  xlim = c(0.7, 6.3), ylim = c(0.4, 3.6))
%lines(y ~ x, data = d[-c(1, 3, 7, 8),])
%lines(y ~ x, data = d[c(1, 3, 5),])
%lines(y ~ x, data = d[c(6, 7, 9),])
%lines(y ~ x, data = d[c(6, 8, 9),])
%points(y ~ x, data = d[c(1, 2, 7, 8), ], pch = 21, cex = 12,
%       bg = c("lightgray", "#F5CF8B", "lightgray", "lightgray"))
%points(y ~ x, data = d[c(3, 4, 5, 6, 9), ], pch = 22, cex = 12,
%       bg = ifelse(1:5 == 5, "#F5CF8B", "lightgray"))
%points(y ~ x, data = d[10, ], pch = 23, bg = "white", cex = 12)
%
%text(d$x, d$y, d$lab)
%text(3.5, 2, "white point\n= D65", cex = 0.7, pos = 3)
%text(4.56, 2, "gamma\n(= 2.4)", cex = 0.7, pos = 3)
%@
%\end{center}
%
%\end{frame}

%% ---------------------------------------------------------------------
%\begin{frame}
%\frametitle{Color spaces}
%
%\textbf{Origin of the package:} Convert colors between various three-dimensional representations of color.
%
%\medskip
%
%\textbf{In particular:} From the perceptually-based HCL (Hue-Chroma-Luminance) to standard Red-Green-Blue (sRGB, and corresponding hex codes) space.
%
%\setkeys{Gin}{width=0.8\textwidth}
%\begin{center}
%<<color-spaces, fig = TRUE, width = 9.5, height = 4.5>>=
%par(mar = rep(0, 4))
%
%d <- data.frame(
%  x = c(1, 1, 2, 2, 3, 4, 5, 5, 5, 6),
%  y = c(1, 3, 1, 3, 2, 2, 1, 3, 2, 2),
%  lab = c("polarLAB", "polarLUV\n(= HCL)", "LAB", "LUV", "XYZ", "RGB", "HLS", "HSV", "sRGB", "hex"),
%  stringsAsFactors = FALSE
%)
%
%plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
%  xlim = c(0.7, 6.3), ylim = c(0.4, 3.6))
%lines(y ~ x, data = d[-c(1, 3, 7, 8),])
%lines(y ~ x, data = d[c(1, 3, 5),])
%lines(y ~ x, data = d[c(6, 7, 9),])
%lines(y ~ x, data = d[c(6, 8, 9),])
%points(y ~ x, data = d[c(1, 2, 7, 8), ], pch = 21, bg = "lightgray", cex = 12)
%points(y ~ x, data = d[c(3, 4, 5, 6, 9), ], pch = 22, bg = "lightgray", cex = 12)
%points(y ~ x, data = d[10, ], pch = 23, bg = "white", cex = 12)
%
%text(d$x, d$y, d$lab)
%text(3.5, 2, "white point\n= D65", cex = 0.7, pos = 3)
%text(4.56, 2, "gamma\n(= 2.4)", cex = 0.7, pos = 3)
%@
%\end{center}
%
%\end{frame}

@@

% ---------------------------------------------------------------------
\begin{frame}
\frametitle{HCL vs.~RGB}

@@

% ---------------------------------------------------------------------
\begin{frame}
\frametitle{HCL vs.~RGB}

\setkeys{Gin}{width=1.0\textwidth}

<<specplot-init, include = FALSE>>=
orig <- c("#008a02", "#00cb00", "#82fd00", "#f5f900", "#f7d600",
          "#d28400", "#fb7502", "#a80003", "#740001", "#8c0084")
hcl  <- c("#b0ceb9", "#9fca91", "#a5bc62", "#b3a924", "#c09000",
          "#c87400", "#c85435", "#b9375c", "#a4176e", "#890075")
@

    \begin{minipage}{0.48\textwidth}
<<specplot-orig, echo = FALSE, fig = TRUE, fig.width = 5, fig.height = 4>>=
specplot(orig, rgb = FALSE, main = "Problem")
@
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
<<specplot-hcl, echo = FALSE, fig = TRUE, fig.width = 5, fig.height = 4>>=
specplot(hcl, rgb = FALSE, main = "Solution")
@
    \end{minipage}

@@

\setkeys{Gin}{width=0.44\textwidth}
<<rgb-rainbow, fig = TRUE, width = 5.2, height = 6>>=
specplot(rainbow(95), rgb = TRUE)
@
\hspace*{0.8cm}
\pause
<<hcl-rainbow, fig = TRUE, width = 5.2, height = 6>>=
specplot(rainbow_hcl(99, c = 60, l = 75), rgb = TRUE)
@

@@

\begin{minipage}[t]{0.49\textwidth}
\textbf{Blues 2:} Single hue. Decreasing chroma with increasing luminance.

\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<blues2-spectrum, fig = TRUE, width = 5.2, height = 4.8>>=
specplot(sequential_hcl(7, "Blues 2"))
@
\end{minipage}%
\hfill
\begin{minipage}[t]{0.49\textwidth}
\textbf{Blues 3:} Single hue. Triangular chroma to achieve higher luminance contrast.

\vspace*{0cm}

\setkeys{Gin}{width=\textwidth}
<<blues3-spectrum, fig = TRUE, width = 5.2, height = 4.8>>=
specplot(sequential_hcl(7, "Blues 3"))
@
\end{minipage}

@@

@@

\begin{frame}
\frametitle{Base R}

@@

@@

\begin{frame}
\frametitle{Do it yourself}

    \textbf{Software:} \emph{colorspace}.
    \begin{itemize}
      \item R (mature): \url{http://colorspace.R-Forge.R-project.org/}.
      \item Python (beta): \url{https://github.com/retostauffer/python-colorspace}.
      \item Web (interactive): \url{http://www.hclwizard.org/}.
    \end{itemize}

    \textbf{Take-home messages:}
    \begin{itemize}
      \item Choose colors carefully.
      \item Make areas of interest stand out from background.
      \item Check robustness.
      \item Software helps you.
    \end{itemize}

@@

\begin{frame}\label{lastslide}
\frametitle{References}

@@

\begin{frame}
	\frametitle{Thank you for your interest!}
	\begin{center}
		\includegraphics[width = .6\textwidth]{images/bobross.jpg}
	\end{center}
\end{frame}
